### 2024-11-22 paper/al.tex
Change file name
Minor organize
Minor changes
WIP: write al

\chapter{Algorithmic Specification Language}
\label{ch:Model}
\noindent

\section{Control flow in SpecTec prose}
\label{sec:spectecp}

@@

% control structure
It might be easy to understand how the control flow of Wasm works, if we assume
that a Wasm code is loaded on a memory space and a pc points to the instruction
to execute.
To describe control flow, it introduces a structure named \textit{block} which
consitutes of a instruction sequence.
When branching instruction in the block is performed, pc can be changed to the
starting point of the block or the end of the block.

@@

\newcommand{\seq}[1]{#1^*}

%% Syntax
\section{Syntax of AL}
\label{syntax}

@@

\newpage
%% Semantics
\section{Semantics of AL}
\label{semantics}

@@

>>>>

\chapter{Algorithmic Specification Language}
\label{ch:al}
\noindent

SpecTec introduces AL, an algorithmic specification language, for specifying
the prose notation.
Given that the prose notation in the official specification is written
informally, it is essential to define AL with rigour.
We begin by how the official specification describes Wasm control flow in
\cref{sec:???}
Next, we then define a formal model for Wasm control flow in AL in
\cref{sec:model}.
We then formalize the syntax and semantics of AL in \cref{sec:formal}.
\red{Finally, we ??? in \cref{sec:spectec-prose}.}

\section{Formal Model for WebAssembly Control Flow}
\label{sec:model}

The way the official specification describe Wasm control flow differs b/w the
prose notation and the formal notation.
In the prose notation, it implicitly assumes a program counter, so it
\textit{jumps} to certain point of the code.
\begin{lstlisting}[style=wasm]
(block (br 0) (unreachable) end)
@\hatcode@
(block (br 0) (unreachable) end)
        @\hatcode@
(block (br 0) (unreachable) end)
                                  @\hatcode@
\end{lstlisting}

\begin{lstlisting}[style=wasm]
(i32.const 1) (if (br 0) (unreachable) else (unreachable) end)
@\hatcode@
(i32.const 1) (if (br 0) (unreachable) else (unreachable) end)
                @\hatcode@
(block (br 0) (unreachable) end)
@\hatcode@
(block (br 0) (unreachable) end)
        @\hatcode@
(block (br 0) (unreachable) end)
                                  @\hatcode@
\end{lstlisting}
Official specification explains Wasm control flow in a different manner in the prose notation and the formal notation.
The keyword is "jump" and "execute"

@@

% control structure
A program counter is imp
It might be easy to understand how the control flow of Wasm works, if we assume
that a Wasm code is loaded on a memory space and a pc points to the instruction
to execute.

To describe control flow, it introduces a structure named \textit{block} which
consitutes of a instruction sequence.
When branching instruction in the block is performed, pc can be changed to the
starting point of the block or the end of the block.

@@

\section{Formalization for Algorithmic Specification Language}
\label{sec:formal}

% TODO: remove seq command
\newcommand{\seq}[1]{#1^*}

%% Syntax
\subsection{Syntax}
\label{syntax}

@@

\newpage
%% Semantics
\subsection{Semantics}
\label{semantics}

@@

\section{Prose Notation Specification for WebAssembly}
\label{sec:spectec-prose}

SpecTec translates declarative definitions in the DSL into algorithmic
definitions in AL.
\cref{ex:al-testop} is generated from \cref{ex:dsl-testop}, representing the
algorithmic specification for the \texttt{testop} instruction using AL:
\\
\begin{example}
\begin{lstlisting}[style=al]
@\textbf{Step\_pure/testop}@ nt testop {
    Assert (check_type_of_stack_top(nt))
    Pop (CONST nt c_1)
    Let c = @$\dollarcode$@testop_(nt, testop, c_1)
    Push (CONST I32 c)
}
\end{lstlisting}
  \label{ex:al-testop}
\end{example}
Using AL, SpecTec generates reStructuredText code.
\cref{fig:spectec-testop} shows the generated document for \texttt{testop}
instruction.
The generated document closely resembles the official document except for some
minor notation changes and some missing hyperlinks.
